---
title: "Week 1 Practice Problems"
author: "Wisley Chen" 
output: pdf_document
date: "2023-01-18"
bibliography: tidyverse.citation.bib
csl: apa.csl
---


```{r setup, include=FALSE}
library(tidyverse)
set.seed(93)
knitr::opts_chunk$set (echo = TRUE, eval = FALSE) 
```

# Introduction to RMarkdown
## Setup
To setup Markdown, we first have a setup code chunk. This consists of the following statements:
```{r}
{r setup, parameters} #Remember, in a real statement, 3 back quotes will 
#indicate both the start and end of the code chunk

#We then call upon the packages we need
library(tidyverse) 
library(knitr)

set.seed(14)

#Then, we can set the parameters in which we want the rest of our RMarkdown
#document to follow
opts_chunk$set (echo = TRUE, eval = FALSE, include = FALSE)

#echo controls whether or not code will be displayed
#eval controls whether output of code will be displayed
#include controls both whether code and output will be displayed
```

## Formatting
There are a lot more formatting tools that can be used within RMarkdown, 
but here are some important ones:
```{r}
# Header 1
## Header 2
### Header 3

*italic*
**bold**
[phrase] (link to phrase)
```



# 2.3 Exercises: Simple functions
## 1. Sums all numbers from 1 to 100

```{r}
n <- 100
sum <- n * (n+1) / 2
```

## 2. Sum all numbers from 1 to 1000

```{r}
sum1000 <- 1000* (1000+1) / 2  
```

## 3. Sum all numbers from 1 to 1000 using seq() and sum()
```{r}
n <- 1000
x <- seq(1,n) #The seq() function generates a vector
#seq (starting number, ending number)
sum(x) #Sums everything in the vector
```

## 4. Functions within functions
```{r}
log(sqrt(100))
```

## 5. Funcstions can be applied to vectors as well
```{r}
x <- seq(1,10)
log10(x^10)
exp(log(x, base = 2))
```
\newpage
# 2.5 Exercises: Introduction to working with datasets
## 1. Attributes of the murderse dataset
```{r}
library(dslabs)
data(murders)
str(murders)

#C IS THE CORRECT ANSWER
```
## 2. Column names of murder dataset
```{r}
murders
#state, abb, region, population
```

## 3. Extract state column from murders dataset
```{r}
a <- murders$state
```

## 4. Extract statement column using [] operator 
```{r}
b <- murders[,1] #[rows, columns]
#a single comma means select all rows

identical(a,b) #Should return true 
```

## 5. Find the number of regions using length() and levels()
```{r}
length(levels(murders$region)) #there are 4 regions
```

## 6. Find how many states are in each region
```{r}
c <- table(murders$region) #table function returns the frequency of unique 
#element within the regions column 
```
\newpage
# 2.8 Exercises: Basic vector creation
## All code to answer 2.8 questions:

```{r}
#1------------------------------------------------------
#Use function c to create a vector with average high temperatures

#2------------------------------------------------------
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro","San Juan", "Toronto")

#3------------------------------------------------------
names(city) <- temp

#4------------------------------------------------------
#Access the first 3 cities
city[1:3]

#5------------------------------------------------------
#Access the temp of Paris and San juan
city[c(3,5)]

#6------------------------------------------------------
seq(12,73)

#7------------------------------------------------------
vector <- seq(1,100)

#8-----------------------------------------------------
#Generate a sequence that starts at 6, ends before 55, and goes
#in intervals of 4/7
length(seq (6,55, by = 4/7)) #This list will have 86 numbers

#9-----------------------------------------------------
class(a <- seq(1,10,0.5)) #numeric type

#10-----------------------------------------------------
class(a <- seq(1, 10)) #integer, because whole numbers

#11-----------------------------------------------------
class(a <- -1L) #R will by default assign numeric type
#to force integer type, you have to add an L

#12-----------------------------------------------------
x <- c("1", "3", "5")
class(x) #Its class is currently character

x <- as.numeric(x)

class(x) #x is now numeric type 
```

# Debugging Exercise
``` {r}
#1--------------------------------------------------------
#"9283" + 43. We get the following error code:
#Error in "9283" + 43 : non-numeric argument to binary operator
#we are adding a non-numeric argument to a numeric argument 


#2--------------------------------------------------------
#remember, the sample () function 
#generates random numbers 
#sample (vector, number of elements we are taking from vector, if we take more
#elements than there is stored in vector, should we replace? TRUE - yes, 
#FALSE - no)

n <- 10
sample(c("Y","N"), n, replace = FALSE) 

#Therefore this statement would not work out because we are taking 10 samples 
#From a vector that has 2 elements, while replace = FALSE
#If replace = TRUE, this would work:

sample(c("Y","N"), n, replace = TRUE) 
```
\newpage
# Addtional Practice
``` {r}
library(tidyverse)
library(ggplot2)
data(msleep)

#1. -------------------------------------------------------------------------
#Print the attributes of the msleep dataset
#a) 
str(msleep)

#b)
msleep[5,] #First 5 rows
msleep[,c("sleep_total","awake")] #sleep_total and awake columns

#c)
odd <- msleep[seq (1, nrow(msleep), 2),] #Displays the even rows
lastfour <- msleep [, (ncol(msleep)-3):ncol(msleep)] #Displays the last 4 rows 
#2. ------------------------------------------------------------------
mean(msleep$brainwt, na.rm = TRUE) #na.rm defines a logical expression 
#indicating whether NA values should be stripped

#3 ------------------------------------------------------------------
factor <- as.factor(msleep$vore)

#4 ------------------------------------------------------------------
sum(as.logical(msleep$sleep_cycle == "NA"))


sum(is.na(msleep$sleep_cycle)) #is.na() function returns a logical vector
#returns TRUE for NA values 
#Therefore the whole expression returns the amount of NA values

#5 ------------------------------------------------------------------
#Create a vector with even numbers from 2 to 50 
even <- seq(2,50, by = 2)
even
#a)
mean(even)

#b) Include 1 and all indices that are a multiple of 3
multiple <- c(even[1], even[seq(3,length(even),by = 3)])
```
## Citations Example (@imai_quantitative_2022)


